---
title: "Data summaries"
output:
  html_document: default
  word_document:
    fig_height: 8
    fig_width: 8
    reference_docx: mystyles.docx
---

```{r setup, echo=FALSE, message=FALSE}

library(dplyr)
library(ggplot2)

DAT.sites <- read.csv("data/sites.csv")
DAT.predigest <- read.csv("data/predigest.csv")
DAT.digest <- read.csv("data/digest.csv")

PointSize <- 3
TextSize <- 20

GOPTS <- theme_bw() +
  theme(text = element_text(size = TextSize),
        legend.position = "bottom",
        legend.key = element_blank())

ShowCode <- FALSE

dat <- DAT.sites %>%
  left_join(DAT.predigest ) %>%
  left_join(DAT.digest)

```

## Data coverage of fire type and time since fire

```{r data_coverage, echo=ShowCode}

ggplot(data = DAT.sites, 
       aes(x = fireType, y = tsf.years)) +
  
  geom_point(position=position_jitter(width=0.2, height=0), 
             size=3) + 
  
  labs(x="", y = "time since fire (years)") +
  
  GOPTS

```

##### page break
## Total carbon

Trend lines are local non-parametric regressions (LOESS) for each fire type, using robust fitting to down-weight outliers.

```{r init_carbon, echo=ShowCode}

ggplot(data = dat, aes(x = tsf.years, y = percentC, colour=fireType)) +
  geom_point(shape = 16, size = PointSize, alpha = 0.5) +
  
  geom_smooth(method = "loess", se = TRUE, 
              method.args = list(family = "symmetric")) +
  
  labs(x = "time since fire (years)", y = "total %C") +
  
  scale_colour_discrete(name="Fire type") +
  
  facet_grid(depth ~ microsite) +
  
  GOPTS

```

##### page break

```{r init_carbon_combined, echo=ShowCode}

ggplot(data = dat, aes(x = tsf.years, y = percentC, colour=fireType)) +
  geom_point(shape = 16, size = PointSize, alpha = 0.5) +
  
  geom_smooth(method = "loess", se = TRUE, 
              method.args = list(family = "symmetric")) +
  
  labs(x = "time since fire (years)", y = "total %C") +
  
  scale_colour_discrete(name="Fire type") +
  
  facet_wrap(~ depth) +
  
  GOPTS

```

##### page break
## Total nitrogen

Trend lines are local non-parametric regressions (LOESS) for each fire type, using robust fitting to down-weight outliers.

```{r init_nitrogen, echo=ShowCode}

ggplot(data = dat, aes(x = tsf.years, y = percentN, colour=fireType)) +
  geom_point(shape = 16, size = PointSize, alpha = 0.5) +

  geom_smooth(method = "loess", se = TRUE, 
              method.args = list(family = "symmetric")) +

  labs(x = "time since fire (years)", y = "%N") +
  
  scale_colour_discrete(name="Fire type") +
  
  facet_grid(depth ~ microsite) +
  
  GOPTS

```

Three samples from a single site have unusually high carbon and nitrogen values:
`r knitr::kable(filter(DAT.predigest, percentC > 15), pad=0)`

Robert suggests that these should be retained.

##### page break

```{r init_nitrogen_combined, echo=ShowCode}

ggplot(data = dat, aes(x = tsf.years, y = percentN, colour=fireType)) +
  geom_point(shape = 16, size = PointSize, alpha = 0.5) +

  geom_smooth(method = "loess", se = TRUE, 
              method.args = list(family = "symmetric")) +

  labs(x = "time since fire (years)", y = "%N") +
  
  scale_colour_discrete(name="Fire type") +
  
  facet_wrap(~ depth) +
  
  GOPTS

```

##### page break
## Total carbon vs total nitrogen

```{r c_vs_n, echo=ShowCode}

ggplot(data = dat, aes(x = percentC, y = percentN)) +
  geom_point(shape = 16, size = PointSize, colour = "darkred", alpha = 0.5) +
  
  labs(x = "total %C", y = "total %N") +
  
  scale_x_log10() +
  scale_y_log10() +
  
  facet_grid(depth ~ microsite) +
  
  GOPTS

```


##### page break
## Recalcitrant carbon

Trend lines are local non-parametric regressions (LOESS) for each fire type, using robust fitting to down-weight outliers.

```{r rp_carbon, echo=ShowCode}

ggplot(data = dat, aes(x = tsf.years, y = percentRPC, colour = fireType)) +
  geom_point(shape = 16, size = PointSize, alpha = 0.5) +
  
  geom_smooth(method = "loess", se = TRUE, 
              method.args = list(family = "symmetric")) +
  
  labs(x = "time since fire (years)", y = "%RPC") +
  
  scale_colour_discrete(name="Fire type") +
  
  facet_grid(depth ~ microsite) +
  
  GOPTS

```

##### page break

```{r rp_carbon_combined, echo=ShowCode}

ggplot(data = dat, aes(x = tsf.years, y = percentRPC, colour = fireType)) +
  geom_point(shape = 16, size = PointSize, alpha = 0.5) +
  
  geom_smooth(method = "loess", se = TRUE, 
              method.args = list(family = "symmetric")) +
  
  labs(x = "time since fire (years)", y = "%RPC") +
  
  scale_colour_discrete(name="Fire type") +
  
  facet_wrap(~ depth) +
  
  GOPTS

```


##### page break
## Recalcitrant vs total carbon

Ratio of RP carbon vs total carbon - looking for any suggestion of pattern in relation to stratifying variables.

Trend lines are local non-parametric regressions (LOESS) for each fire type, using robust fitting to down-weight outliers.


```{r rpc_ratio_vs_totalc, echo = ShowCode}

dat <- mutate(dat, ratioRPC = percentRPC / percentC)

YLIM <- c(0, 0.6)

ggplot(data = dat, aes(x = tsf.years, y = ratioRPC, colour = fireType)) +
  geom_point(shape = 16, size = PointSize, alpha = 0.5,
             position = position_jitter(width=0.2, height=0)) +
  
  geom_smooth(method = "loess", se = TRUE, 
              method.args = list(family = "symmetric")) +
  
  scale_y_continuous(limits = YLIM) +

  scale_colour_discrete(name="Fire type") +
  
  labs(x = "time since fire (years)", y = "ratio RPC to total C") +

  facet_grid(depth ~ microsite) +
  
  GOPTS

```

##### page break

```{r rpc_ratio_vs_totalc_combined, echo = ShowCode}

ggplot(data = dat, aes(x = tsf.years, y = ratioRPC, colour = fireType)) +
  geom_point(shape = 16, size = PointSize, alpha = 0.5,
             position = position_jitter(width=0.2, height=0)) +
  
  geom_smooth(method = "loess", se = TRUE, 
              method.args = list(family = "symmetric")) +
  
  scale_y_continuous(limits = YLIM) +

  scale_colour_discrete(name="Fire type") +
  
  labs(x = "time since fire (years)", y = "ratio RPC to total C") +

  facet_grid(~ depth) +
  
  GOPTS

```

##### page break
Another way of looking at this, plotting RP carbon vs total carbon directly.

```{r rpc_vs_totalc, echo=ShowCode}

ggplot(data = dat, aes(x = percentC, y = percentRPC, colour = tsf.years)) +
  geom_point(shape = 0, size = PointSize, alpha = 0.5) +
  
  #scale_x_log10() +
  #scale_y_log10() +
  scale_color_gradient2(name = "Time since fire",
                        low = "red", mid = "lightgrey", high = "darkblue",
                        midpoint = 20) +
  
  labs(x = "% total carbon", y = "% recalcitrant carbon") +

  facet_grid(depth + fireType ~ microsite) +
  
  GOPTS

```

##### page break

```{r rpc_vs_totalc_combined, echo=ShowCode}

ggplot(data = dat, aes(x = percentC, y = percentRPC, colour = tsf.years)) +
  geom_point(shape = 0, size = PointSize, alpha = 0.5) +
  
  #scale_x_log10() +
  #scale_y_log10() +
  scale_color_gradient2(name = "Time since fire",
                        low = "red", mid = "lightgrey", high = "darkblue",
                        midpoint = 20) +
  
  labs(x = "% total carbon", y = "% recalcitrant carbon") +

  facet_grid(fireType ~ depth) +
  
  GOPTS

```

